---
params: 
  title: ""
  publication_date: ""
  doi: ""
output: 
  html_document:
    anchor_sections: false
    theme: null
    highlight: null
    mathjax: null
    css: ["style.css", "https://fonts.googleapis.com/css?family=Source+Sans+Pro:400,700&display=swap"]
    self_contained: true
title: "`r params$title`"
---

```{r general-setup, include=FALSE}
## This file contains the FRENCH version of the data story 

# Set general chunk options
knitr::opts_chunk$set(echo = FALSE, fig.showtext = TRUE, fig.retina = 3, 
                      fig.align = "center", warning = FALSE, message = FALSE)

# Install pacman package if needed
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
}

# Install snf.datastory package if not available, otherwise load it
if (!require("snf.datastory")) {
  if (!require("devtools")) {
    install.packages("devtools")
    library(devtools)
  }
  install_github("snsf-data/snf.datastory")
  library(snf.datastory)
}

# Load packages
p_load(tidyverse,
       lubridate,
       scales, 
       conflicted, 
       jsonlite,
       here, 
       ggiraph)

# Conflict preferences
conflict_prefer("filter", "dplyr")
conflict_prefer("get_datastory_theme", "snf.datastory")
conflict_prefer("get_datastory_scheme", "snf.datastory")

# Increase showtext package font resolution
showtext_opts(dpi = 320)

# Set the locale for date formatting (Windows)
Sys.setlocale("LC_TIME", "French")

# Create function to print number with local language-specific format 
print_num <- function(x) snf.datastory::print_num(x, lang = "fr")

# Knitr hook for local formatting of printed numbers
knitr::knit_hooks$set(
  inline <- function(x) {
    if (!is.numeric(x)) {
      x
    } else {
      print_num(x)
    }
  }
)

# Function that returns the desired language version of a string, given three
# language strings are provided to the function 
translate <- function(en_string = "NOT TRANSLATED",
                      de_string = "NOT TRANSLATED",
                      fr_string = "NOT TRANSLATED",
                      lang = "en") {
  string <- case_when(lang == "en" ~ en_string, 
                      lang == "de" ~ de_string, 
                      lang == "fr" ~ fr_string, 
                      TRUE ~ "NO VALID LANGUAGE SPECIFIED")
  string
}
```

```{r print-header-infos, results='asis'}
# Add publication date to header
cat(format(as_datetime(params$publication_date), "%d.%m.%Y"))

# Register the Google font (same as Data Portal, is not loaded twice)
cat(paste0("<link href='https://fonts.googleapis.com/css?family=", 
           "Source+Sans+Pro:400,700&display=swap' rel='stylesheet'>"))
```

```{r story-specific-setup, include=FALSE}
# Load data about funding in 2021
funding_in_2021 <- read_csv(here("data", "prelim_funding_in_2021_20220131.csv"))
```


<!-- Short lead (2-3 sentences) in bold -->

__Premiers chiffres provisoires concernant le soutien financier du FNS en 2021 : nous avons investi 882 millions de francs dans de nouveaux projets, des bourses et des publications.__

La moitié des fonds est consacrée à l’encouragement de projets, qui permet à des scientifiques expérimentés de réaliser leurs idées sur des thèmes de leur choix. En outre, le soutien à la carrière des jeunes chercheuses et chercheurs engage un quart du budget, et le travail des groupes de recherche dans les programmes environ 20 pour cent. Le reste des ressources se répartit entre les infrastructures et la communication scientifique. Fin 2021, quelque 5700 projets financés par le FNS étaient en cours, auxquels participent 20000 chercheuses et chercheurs dans les hautes écoles et d’autres institutions. La plupart des projets courent sur plusieurs années.


<div class="hide-mobile hide-tablet" style="width: 800px;position: relative;max-width: 800px;left: -70px;border-top: 1px solid #d3d3d3;border-bottom: 1px solid #d3d3d3;padding-top: 20px;margin: 20px 0;">
<div class="plot-box">
<div class="plot-title">Montants des encouragements FNS approuvés en 2021 (chiffres provisoires)</div>

```{r, out.width="100%", fig.height=2.5}
draw_funding_portfolio_5 <- function(lang = "en") {
  # Mapping table for the translation of the Level 1 Funding Instruments
  mapping_fi <- tibble(
    en = c("Project funding", "Careers", "Programmes", "Infrastructure", 
           "Science communication"),
    de = c("Projekte", "Karrieren", "Programme", "Infrastrukturen", 
           "Wissenschaftskommunikation"),
    fr = c("Encouragement de projets", "Carrières", "Programmes", 
           "Infrastructures", "Communication scientifique")
  ) %>% 
    # Now choose the English string (to join and the destination lang string)
    mutate(FundingInstrumentGaLevel1 = en) %>% 
    select(FundingInstrumentGaLevel1, destination = any_of(lang))

  # Compile the text string to be displayed in the tooltip
  tooltip_text <- funding_in_2021 %>% 
    group_by(FundingInstrumentGaLevel1, FundingInstrumentGaReporting_Fr) %>%
    summarise(scheme_granted = sum(AmountGranted, na.rm = TRUE),
              n = n()) %>% 
    ungroup() %>% 
    group_by(FundingInstrumentGaLevel1) %>% 
    arrange(desc(scheme_granted)) %>% 
    summarise(tooltip = paste0("<strong>", FundingInstrumentGaReporting_Fr, 
                               "</strong> (", 
                               print_num(round(scheme_granted / 1000000, 
                                               digits = 1)), 
                               " mio CHF approuvés en ", print_num(n), 
                               " contributions)",
                               collapse = "<br>"))
  
  
  # Prepare plot data, calculate frequencies of granted amounts per LVL 1 FI
  # (Careers, Programmes...)
  plot_data <- funding_in_2021 %>% 
    group_by(FundingInstrumentGaLevel1) %>%
    summarise(lvl_1_granted = sum(AmountGranted, na.rm = TRUE),
              n = n()) %>%
    # Get translated version of FundingInstrumentGaLevel1
    left_join(mapping_fi, by = "FundingInstrumentGaLevel1") %>% 
    # Add scheme-level tooltip text data
    left_join(tooltip_text, by = "FundingInstrumentGaLevel1") %>% 
    # Overwrite with the destination language string
    mutate(FundingInstrumentGaLevel1 = destination) %>% 
    mutate(lvl_1_freq = lvl_1_granted / sum(lvl_1_granted)) %>%
    # Order the data for the plot: by LVL 1 share size
    arrange(desc(lvl_1_freq)) %>%
    mutate(order = row_number()) %>%
    # Now calculate where the segment has to start (and end) per LVL 1 FI
    mutate(seg_start = map_dbl(order, function(x) {
      prev_dist <- filter(., order < x) %>%
        pull(lvl_1_freq) %>%
        sum(na.rm = TRUE)
      prev_dist
    }),
    seg_end = seg_start + lvl_1_freq)
  
  # Color settings: Colors for the LVL 1 instruments are defined as in the
  # Data Portal (YAAY scheme)
  color_palette_lvl1 <-
    tibble(
      FundingInstrumentGaLevel1 = c(
        translate("Project funding", "Projekte", "Encouragement de projets", 
                  lang),
        translate("Careers", "Karrieren", "Carrières", lang),
        translate("Programmes", "Programme", "Programmes", lang),
        translate("Infrastructure", "Infrastrukturen", "Infrastructures", lang),
        translate("Science communication", "Wissenschaftskommunikation", 
                  "Communication scientifique", lang)
      ),
      # Use slightly adapted SNSF colors for this kind of plot
      color_lvl1 = c("#3D7D9F", 
                     "#83D0F5",
                     "#FBBE5E",
                     "#71B294",
                     "#9D90B9")
    )
  
  # Create the plot
  p_funding_portfolio_5 <- plot_data %>%
    ggplot(aes(x = 1, y = lvl_1_freq ,
               fill = reorder(FundingInstrumentGaLevel1, -order), 
               tooltip = tooltip)) +
    geom_col_interactive(color = "white", width = 0.7) +
    # Level 1 FI labels
    # Bold text (FI and percentage)
    geom_text(aes(x = ifelse(lvl_1_freq < 0.15, 1.375, 1.65),
                  y = seg_start,
                  label = paste0(FundingInstrumentGaLevel1, " (",
                                 round(lvl_1_freq  * 100), "%)"),
                  angle = ifelse(lvl_1_freq < 0.15, 90, 0),
                  hjust = 0),
              vjust = 1,
              color = "#4d4d4d", family = "Source Sans Pro",
              size = 2.75, fontface = "bold") +
    # Plain text (amount)
    geom_text(aes(x = ifelse(lvl_1_freq < 0.15, 1.375, 1.5),
                  y = ifelse(lvl_1_freq < 0.15, seg_start + 0.018, seg_start),
                  label = paste0(round((lvl_1_granted / 1000000)),
                                 translate(" million CHF", 
                                           " Mio. CHF", 
                                           " mio CHF", 
                                           lang)
                  ),
                  angle = ifelse(lvl_1_freq < 0.15, 90, 0)),
              vjust = 1,
              hjust = 0,
              color = "#4d4d4d", family = "Source Sans Pro",
              size = 2.75, fontface = "plain") +
    coord_flip() +
    get_datastory_theme(remove_plot_margin = TRUE) + 
    scale_fill_manual(
      values = setNames(color_palette_lvl1$color_lvl1,
                        color_palette_lvl1$FundingInstrumentGaLevel1),
      guide = "none") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0.01, 0.01),
                       labels = percent_format(accuracy = 1)) +
    scale_color_manual(
      guide = "none",
      values = setNames(color_palette_lvl1$color_lvl1,
                        color_palette_lvl1$FundingInstrumentGaLevel1)
    ) +
    labs(x = NULL, y = NULL) +
    theme(axis.text.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x = element_line(color = "#d3d3d3", size = 0.3)) +
    expand_limits(x = 3.5)
  p_funding_portfolio_5
}

# Create ggiraph object
girafe(ggobj = draw_funding_portfolio_5("fr"), 
       height_svg = 2.5, 
       options = list(
         opts_toolbar(saveaspng = FALSE),
         opts_hover(css = "fill:#F08262;stroke:#F08262;"),
         opts_tooltip(
           css = get_ggiraph_tooltip_css(),
           opacity = 0.8,
           delay_mouseover = 0,
           delay_mouseout = 0
         )
       ))
```
<div class="caption">
Données: <a href="https://github.com/snsf-data/datastory_preliminary_funding_in_2021/tree/main/data/prelim_funding_in_2021_20220131.csv" target="_blank">Montants des encouragements FNS approuvés en 2021 (chiffres provisoires, état au 31 janvier 2022)</a>.
</div>
</div>
</div>

<div class="hide-desktop">
<div class="plot-box">
<div class="plot-title">Montants des encouragements FNS approuvés en 2021 (chiffres provisoires)</div>
```{r, out.width="100%", fig.width=3, fig.height=6}
draw_funding_portfolio_5_mob <- function(lang = "en") {
  # Mapping table for the translation of the Level 1 Funding Instruments
  mapping_fi <- tibble(
    en = c("Project funding", "Careers", "Programmes", "Infrastructure", 
           "Science communication"),
    de = c("Projekte", "Karrieren", "Programme", "Infrastrukturen", 
           "Wissenschaftskommunikation"),
    fr = c("Encouragement de projets", "Carrières", "Programmes", 
           "Infrastructures", "Communication scientifique")
  ) %>% 
    # Now choose the English string (to join and the destination lang string)
    mutate(FundingInstrumentGaLevel1 = en) %>% 
    select(FundingInstrumentGaLevel1, destination = any_of(lang))

  # Compile the text string to be displayed in the tooltip
  tooltip_text <- funding_in_2021 %>% 
    group_by(FundingInstrumentGaLevel1, FundingInstrumentGaReporting_Fr) %>%
    summarise(scheme_granted = sum(AmountGranted, na.rm = TRUE),
              n = n()) %>% 
    ungroup() %>% 
    group_by(FundingInstrumentGaLevel1) %>% 
    arrange(desc(scheme_granted)) %>% 
    summarise(tooltip = paste0("<strong>", FundingInstrumentGaReporting_Fr, 
                               "</strong> (", 
                               print_num(round(scheme_granted / 1000000, 
                                               digits = 1)), 
                               " mio CHF approuvés en ", print_num(n), 
                               " contributions)",
                               collapse = "<br>"))
  
  
  # Prepare plot data, calculate frequencies of granted amounts per LVL 1 FI
  # (Careers, Programmes...)
  plot_data <- funding_in_2021 %>% 
    group_by(FundingInstrumentGaLevel1) %>%
    summarise(lvl_1_granted = sum(AmountGranted, na.rm = TRUE),
              n = n()) %>%
    # Get translated version of FundingInstrumentGaLevel1
    left_join(mapping_fi, by = "FundingInstrumentGaLevel1") %>% 
    # Add scheme-level tooltip text data
    left_join(tooltip_text, by = "FundingInstrumentGaLevel1") %>% 
    # Overwrite with the destination language string
    mutate(FundingInstrumentGaLevel1 = destination) %>% 
    mutate(lvl_1_freq = lvl_1_granted / sum(lvl_1_granted)) %>%
    # Order the data for the plot: by LVL 1 share size
    arrange(desc(lvl_1_freq)) %>%
    mutate(order = nrow(.) - row_number()) %>%
    # Now calculate where the segment has to start (and end) per LVL 1 FI
    mutate(seg_start = map_dbl(order, function(x) {
      prev_dist <- filter(., order < x) %>%
        pull(lvl_1_freq) %>%
        sum(na.rm = TRUE)
      prev_dist
    }),
    seg_end = seg_start + lvl_1_freq)
  
  # Color settings: Colors for the LVL 1 instruments are defined as in the
  # Data Portal (YAAY scheme)
  color_palette_lvl1 <-
    tibble(
      FundingInstrumentGaLevel1 = c(
        translate("Project funding", "Projekte", "Encouragement de projets", 
                  lang),
        translate("Careers", "Karrieren", "Carrières", lang),
        translate("Programmes", "Programme", "Programmes", lang),
        translate("Infrastructure", "Infrastrukturen", "Infrastructures", lang),
        translate("Science communication", "Wissenschaftskommunikation", 
                  "Communication scientifique", lang)
      ),
      # Use slightly adapted SNSF colors for this kind of plot
      color_lvl1 = c("#3D7D9F", 
                     "#83D0F5",
                     "#FBBE5E",
                     "#71B294",
                     "#9D90B9")
    )
  
  # Create the plot
  p_funding_portfolio_5 <- plot_data %>%
    ggplot(aes(x = 1, y = lvl_1_freq ,
               fill = reorder(FundingInstrumentGaLevel1, -order), 
               tooltip = tooltip)) +
    geom_col_interactive(color = "white", width = 0.7) +
    # Level 1 FI labels
    # Bold text (FI and percentage)
    geom_text(aes(x = 1.375,
                  y = seg_end,
                  label = paste0(FundingInstrumentGaLevel1, " (",
                                 round(lvl_1_freq  * 100), "%)"),
                  # angle = ifelse(lvl_1_freq < 0.15, 90, 0),
                  hjust = 0),
              vjust = 1,
              color = "#4d4d4d", family = "Source Sans Pro",
              size = 3, fontface = "bold") +
    # Plain text (amount)
    geom_text(aes(x = 1.375,
                  y = seg_end - 0.02,
                  label = paste0(round((lvl_1_granted / 1000000)),
                                 translate(" million CHF", 
                                           " Mio. CHF", 
                                           " mio CHF", 
                                           lang)
                  )#,
                  # angle = ifelse(lvl_1_freq < 0.15, 90, 0)
                  ),
              vjust = 1,
              hjust = 0,
              color = "#4d4d4d", family = "Source Sans Pro",
              size = 3, fontface = "plain") +
    # coord_flip() +
    get_datastory_theme() + 
    scale_fill_manual(
      values = setNames(color_palette_lvl1$color_lvl1,
                        color_palette_lvl1$FundingInstrumentGaLevel1),
      guide = "none") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0.01, 0.01),
                       labels = percent_format(accuracy = 1)) +
    scale_color_manual(
      guide = "none",
      values = setNames(color_palette_lvl1$color_lvl1,
                        color_palette_lvl1$FundingInstrumentGaLevel1)
    ) +
    labs(x = NULL, y = NULL) +
    theme(axis.text.x = element_blank(),
          panel.grid.major.y = element_blank()) +
    expand_limits(x = 4)
  p_funding_portfolio_5
}

# Create ggiraph object
girafe(ggobj = draw_funding_portfolio_5_mob("fr"), 
       width_svg = 3,
       height_svg = 6, 
       options = list(
         opts_toolbar(saveaspng = FALSE),
         opts_hover(css = "fill:#F08262;stroke:#F08262;"),
         opts_tooltip(
           css = get_ggiraph_tooltip_css(),
           opacity = 0.8,
           delay_mouseover = 0,
           delay_mouseout = 0
         )
       ))
```
<div class="caption">
Données: <a href="https://github.com/snsf-data/datastory_preliminary_funding_in_2021/tree/main/data/prelim_funding_in_2021_20220131.csv" target="_blank">Montants des encouragements FNS approuvés en 2021 (chiffres provisoires, état au 31 janvier 2022)</a>.
</div>
</div>
</div>

Les données et le code de ce récit de données sont <a href="https://github.com/snsf-data/datastory_preliminary_funding_in_2021" target="_blank">disponibles sur Github</a> et <a href="https://doi.org/10.5281/zenodo.6022479" target="_blank">archivés sur Zenodo</a>.